---
title: "Intro and descriptives"
title-block-banner: "#009ca6"
title-block-banner-color: "#FFFFFF"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://www.ri.se/en/kbm
  orcid: 0000-0003-1669-592X
date: last-modified
date-format: iso
always_allow_html: true
format: html
execute:
  echo: true
  warning: false
  message: false
  cache: false
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r}
#| label: setup

# one package below requires that you use devtools to install it manually:
# first install devtools by
# install.packages('devtools')

library(RISEkbmRasch) # devtools::install_github("pgmj/RISEkbmRasch")
library(grateful)
library(ggrepel)
library(car)
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(mirt)
library(psych)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(cowplot)
library(formattable) 
library(glue)

### optional libraries
#library(TAM)
#library(skimr)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

source("RISE_theme.R")
```


```{r}
### import data
df <- read_excel("data/MasterFile_CLEAN.xlsx") %>% 
  rename(gender = WhatisyourgenderSelectedChoice,
         age = Whatisyourage,
         citizenship = WhatisyourcitizenshipSelectedChoice) %>% 
  select(gender,age,citizenship,starts_with("mhc")) %>% 
  select(-starts_with("mhc_"))

### Load item information
# make sure that variable names in df match with itemlabels$itemnr
itemlabels <- read_excel("data/itemlabels_MHC_SF.xlsx") %>% 
  mutate(item = str_squish(item))

#---- Create DIF variables----
  
# DIF variables into vectors, recoded as factors since DIF functions need this
dif.gender <- as.factor(df$gender)
dif.age <- df$age
dif.citizenship <- as.factor(df$citizenship)
# then remove them from dataframe, since we need a dataframe with only item data for the Rasch analyses
df$gender <- NULL
df$age <- NULL
df$citizenship <- NULL

```

```{r}
#| eval: false
### other, bigger dataset from [@echeverría2017]
### read file
df <- read_excel("data/data_sMHCSF_Echeverria2017.xlsx")

### remove NA
df.omit.na <- df %>% na.omit()

### create dif variables
dif.sex <- factor(df.omit.na$Sex)
df.omit.na$Sex <- NULL

itemlabels <- data.frame(
  itemnr = paste0("mhc_",c(1:14)),
  item = names(df.omit.na)
)

names(df.omit.na) <- itemlabels$itemnr

df.omit.na <- df.omit.na %>% 
  mutate(across(everything(), ~ car::recode(.x,"'Never'=0;'1 or 2 times a month'=1;'About 1 time a week'=2;'About 2 or 3 times a week'=3;'Almost daily'=4;'Daily'=5", as.factor = FALSE)))

```


## Background

The Mental Health Continuum-Short Form (MHC-SF) is a well-being scale designed to assess emotional, psychological and social well-being [@keyes2009]. It is derived from the long form (MHC-LF), which is based on dimensions of Ryff's [@ryff1989] model of psychological well-being and Keyes' [@keyes1998] model of social well-being. 

https://doi.org/10.6084/m9.figshare.19174205.v2


### Items
```{r}
RIlistitems(df)
```

### Response categories

- 'Never', 
- '1 or 2 times a month', 
- 'About 1 time a week', 
- 'About 2 or 3 times a week', 
- 'Almost daily', 
- 'Daily'

## Demographics
```{r}
#| layout-ncol: 2
RIdemographics(dif.gender, "Gender")
RIdemographics(dif.citizenship, "Citizenship")
```

```{r}
hist(dif.age, main = "Age distribution", xlab = "Age", col = "#009ca6", border = "white",breaks = 20)
abline(v = mean(dif.age, na.rm = TRUE), col = "red", lwd = 2, lty = 2)
```

## Descriptives

Response distribution for all items are summarized below.

```{r}
#| tbl-cap: "Total number of responses for all items"
# Specify the numeric-to-label mapping
numeric_to_labels <- c('Never', '1 or 2 times a month', 'About 1 time a week', 'About 2 or 3 times a week', 'Almost daily', 'Daily')
df %>% 
  mutate(across(everything(), ~ factor(.x, labels = numeric_to_labels))) %>% 
  RIallresp()
```

### Descriptives - item level

```{r}
#| column: margin
RIlistItemsMargin(df, fontsize = 12)
```

::: panel-tabset
#### Tile plot
```{r}
RItileplot(df)
```
#### Stacked bars
```{r}
RIbarstack(df)
```
#### Barplots
```{r}
#| layout-ncol: 2
RIbarplot(df)
```
#### Missing responses
```{r}
RImissing(df)
```
:::

No missing data. Some items have a low number of responses (< 10) in the bottom categories, which impacts the reliability of the analysis of those categories.

## References
