---
title: "Intro and descriptives"
title-block-banner: "#009ca6"
title-block-banner-color: "#FFFFFF"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://www.ri.se/en/kbm
  orcid: 0000-0003-1669-592X
date: last-modified
date-format: iso
always_allow_html: true
format: html
execute:
  echo: true
  warning: false
  message: false
  cache: false
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
bibliography: references.bib
---

```{r}
#| label: setup

# one package below requires that you use devtools to install it manually:
# first install devtools by
# install.packages('devtools')

library(RISEkbmRasch) # devtools::install_github("pgmj/RISEkbmRasch")
library(grateful)
library(ggrepel)
library(car)
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(mirt)
library(psych)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(patchwork)
library(formattable) 
library(glue)

### optional libraries
#library(TAM)
#library(skimr)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

source("RISE_theme.R")
```

```{r}
### other, bigger dataset from [@echeverría2017]
### read file
df.all <- read_excel("data/data_sMHCSF_Echeverria2017.xlsx")
df <- df.all
### create dif variables
dif.sex <- factor(df$Sex)
df$Sex <- NULL

### Load item information
# make sure that variable names in df match with itemlabels$itemnr
itemlabels <- read_excel("data/itemlabels_MHC_SF.xlsx") %>% 
  mutate(item = str_squish(item))

names(df) <- itemlabels$itemnr

df <- df %>%
  mutate(across(everything(), ~ car::recode(.x,"'Never'=0;'1 or 2 times a month'=1;'About 1 time a week'=2;'About 2 or 3 times a week'=3;'Almost daily'=4;'Daily'=5", as.factor = FALSE)))
```

## Introduction

In a paper recently made available as a preprint [@johansson], we argue
that the basic aspects of a psychometric analysis should include
information about five basic criteria:

-   Unidimensionality
-   Response categories
-   Invariance
-   Targeting
-   Measurement uncertainties (reliability)

## About the MHC-SF

The Mental Health Continuum-Short Form (MHC-SF) is a well-being scale
designed to assess emotional, psychological and social well-being
[@keyes2009]. The dataset for this analysis was made available as open
data connected to a psychometric analysis publication [@echeverría2017;
@echeverría].

While there are papers suggesting a three factor or bi-factor model for the MHC-SF, in practice it is usually used as a sum scored unidimensional scale, which is why our analyses will focus on how the data fits a unidimensional model.

Here is some text about the MHC-SF that I wrote some years back when investigating different well-being questionnaires:

> Flourishing & Mental Health Continuum Short Form (MHC-SF, Keyes, 2014). 14 items (derived from the 40-item long version) measuring the following three aspects of flourishing (Keyes, 2002, 2005, 2006): emotional wellbeing (3 items; e.g., “In the past month, how often have you felt happy?”); psychological wellbeing (6 items); e.g., “In the past month, how often did you feel good at managing the responsibilities of your daily life?”); and social wellbeing (5 items; e.g., “In the past month, how often did you feel that you belonged to a community like a social group, your school, or your neighborhood?”). Participants rated their responses using a scale with the following labels: 1 (Never), 2 (Once or twice), 3 (About once a week), 4 (Two or three times a week), 5 (Almost every day), and 6 (Every day). (Paragraph copied from (Sahdra et al., 2016)).

> Other factor structures have been proposed as well, such as a bi-factor model, combining the three factors with a general well-being factor (de Bruin & du Plessis, 2015; Lamborn et al., 2018) and a 2-factor model with mental health and mental illness as correlated factors (Lamers et al., 2011), but in general the 3-factor model seems replicable. IRT analysis (Lamers et al., 2012) showed DIF for gender (item 8), age (items 10 and 12) and education level (item 5), estimated to be of non-significant impact on latent scales (but says nothing about sum scores). Items 4 and 5 from the social well-being subscale have been found to have very low factors loadings (.32-.34) in two papers (Lamers et al., 2011; Lupano Perugini et al., 2017).

### Items

```{r}
RIlistitems(df)
```

### Response categories

-   'Never',
-   '1 or 2 times a month',
-   'About 1 time a week',
-   'About 2 or 3 times a week',
-   'Almost daily',
-   'Daily'

## Demographics

```{r}
#| layout-ncol: 2
RIdemographics(dif.sex, "Gender")
```

## Descriptives

Response distribution for all items are summarized below.

```{r}
#| tbl-cap: "Total number of responses for all items"
# Specify the numeric-to-label mapping
numeric_to_labels <- c('Never', '1 or 2 times a month', 'About 1 time a week', 'About 2 or 3 times a week', 'Almost daily', 'Daily')
df %>% 
  mutate(across(everything(), ~ factor(.x, labels = numeric_to_labels))) %>% 
  RIallresp()
```

### Descriptives - item level

```{r}
#| column: margin
RIlistItemsMargin(df, fontsize = 12)
```

::: panel-tabset
#### Tile plot

```{r}
RItileplot(df)
```

#### Stacked bars

```{r}
RIbarstack(df) +
  scale_fill_viridis_d(labels = rev(numeric_to_labels), direction = -1)
```

#### Barplots

```{r}
#| layout-ncol: 2
RIbarplot(df)
```

#### Missing responses

```{r}
RImissing(df)
```
:::

## References
