---
title: "Intro and descriptives"
title-block-banner: "#009ca6"
title-block-banner-color: "#FFFFFF"
author: 
  name: Magnus Johansson
  affiliation: RISE Research Institutes of Sweden
  affiliation-url: https://www.ri.se/en/kbm
  orcid: 0000-0003-1669-592X
date: last-modified
date-format: iso
always_allow_html: true
format: html
execute:
  echo: true
  warning: false
  message: false
  cache: false
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
bibliography: references.bib
---

```{r}
#| label: setup

# one package below requires that you use devtools to install it manually:
# first install devtools by
# install.packages('devtools')

library(RISEkbmRasch) # devtools::install_github("pgmj/RISEkbmRasch")
library(grateful)
library(ggrepel)
library(car)
library(kableExtra)
library(readxl)
library(tidyverse)
library(eRm)
library(mirt)
library(psych)
library(psychotree)
library(matrixStats)
library(reshape)
library(knitr)
library(patchwork)
library(formattable) 
library(glue)

### optional libraries
#library(TAM)
#library(skimr)

### some commands exist in multiple packages, here we define preferred ones that are frequently used
select <- dplyr::select
count <- dplyr::count
recode <- car::recode
rename <- dplyr::rename

source("RISE_theme.R")
```

```{r}
### import data https://doi.org/10.6084/m9.figshare.19174205.v2
# df <- read_excel("data/MasterFile_CLEAN.xlsx") %>% 
#   rename(gender = WhatisyourgenderSelectedChoice,
#          age = Whatisyourage,
#          citizenship = WhatisyourcitizenshipSelectedChoice) %>% 
#   select(gender,age,citizenship,starts_with("mhc")) %>% 
#   select(-starts_with("mhc_"))
# 
# ### Load item information
# # make sure that variable names in df match with itemlabels$itemnr
# itemlabels <- read_excel("data/itemlabels_MHC_SF.xlsx") %>% 
#   mutate(item = str_squish(item))
# 
# #---- Create DIF variables----
#   
# # DIF variables into vectors, recoded as factors since DIF functions need this
# dif.gender <- as.factor(df$gender)
# dif.age <- df$age
# dif.citizenship <- as.factor(df$citizenship)
# # then remove them from dataframe, since we need a dataframe with only item data for the Rasch analyses
# df$gender <- NULL
# df$age <- NULL
# df$citizenship <- NULL

```

```{r}
### other, bigger dataset from [@echeverría2017]
### read file
df.all <- read_excel("data/data_sMHCSF_Echeverria2017.xlsx")
df <- df.all
### create dif variables
dif.sex <- factor(df$Sex)
df$Sex <- NULL

### Load item information
# make sure that variable names in df match with itemlabels$itemnr
itemlabels <- read_excel("data/itemlabels_MHC_SF.xlsx") %>% 
  mutate(item = str_squish(item))

names(df) <- itemlabels$itemnr

df <- df %>%
  mutate(across(everything(), ~ car::recode(.x,"'Never'=0;'1 or 2 times a month'=1;'About 1 time a week'=2;'About 2 or 3 times a week'=3;'Almost daily'=4;'Daily'=5", as.factor = FALSE)))
```

## Introduction

In a paper recently made available as a preprint [@johansson], we argue
that the basic aspects of a psychometric analysis should include
information about five basic criteria:

-   Unidimensionality
-   Response categories
-   Invariance
-   Targeting
-   Measurement uncertainties (reliability)

## About the MHC-SF

The Mental Health Continuum-Short Form (MHC-SF) is a well-being scale
designed to assess emotional, psychological and social well-being
[@keyes2009]. The dataset for this analysis was made available as open
data connected to a psychometric analysis publication [@echeverría2017;
@echeverría].

While there are papers suggesting a three factor or bi-factor model for the MHC-SF, in practice it is usually used as a sum scored unidimensional scale, which is why our analyses will focus on how the data fits a unidimensional model.

### Items

```{r}
RIlistitems(df)
```

### Response categories

-   'Never',
-   '1 or 2 times a month',
-   'About 1 time a week',
-   'About 2 or 3 times a week',
-   'Almost daily',
-   'Daily'

## Demographics

```{r}
#| layout-ncol: 2
RIdemographics(dif.sex, "Gender")
```

## Descriptives

Response distribution for all items are summarized below.

```{r}
#| tbl-cap: "Total number of responses for all items"
# Specify the numeric-to-label mapping
numeric_to_labels <- c('Never', '1 or 2 times a month', 'About 1 time a week', 'About 2 or 3 times a week', 'Almost daily', 'Daily')
df %>% 
  mutate(across(everything(), ~ factor(.x, labels = numeric_to_labels))) %>% 
  RIallresp()
```

### Descriptives - item level

```{r}
#| column: margin
RIlistItemsMargin(df, fontsize = 12)
```

::: panel-tabset
#### Tile plot

```{r}
RItileplot(df)
```

#### Stacked bars

```{r}
RIbarstack(df) +
  scale_fill_viridis_d(labels = rev(numeric_to_labels), direction = -1)
```

#### Barplots

```{r}
#| layout-ncol: 2
RIbarplot(df)
```

#### Missing responses

```{r}
RImissing(df)
```
:::

## References
